---
title: "BI612 Homework 1"
author: "Caitlin Sutherland"
date: "12/01/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Part I - Working with vectors and data frames

**Question 1. Create a new vector in R with the data for each strain.**

```{r}
a=c(0.17,0.20,0.13,0.20,0.24,0.25,0.15,0.21,0.21)
b=c(0.47,0.21,0.34,0.32,0.33,0.29,0.18,0.20,0.25)
c=c(0.44,0.33,0.37,0.33,0.31,0.27,0.22,0.30,0.28)
```

**Question 2. What is the difference between the means of the two transgenic strains?**

```{r}
'MeanA' = mean(a)
'MeanB' = mean(b)
'DifferenceAB' = MeanB-MeanA
```

The mean of transgenic strain A is `r MeanA` and the mean of transgenic strain B is `r MeanB`. The difference between the two means is `r DifferenceAB`.

**Question 3. What is the difference between the means of the common strain and transgenic strain A?**

```{r}
'MeanCommon' = mean(c)
'DifferenceAC' = MeanCommon-MeanA
```

The mean of the common strain is `r MeanCommon` and the mean of transgenic strain A is `r MeanA`. The difference between the two means is `r DifferenceAC`. 

**Question 4. Create a histogram of the hydrogen peroxide concentrations in the common strain.**

```{r}
hist(c,main=paste("Histogram of common strain"),xlab="Hydrogen peroxide concentration (µmoles/g)")
```

**Question 5. Enter each column as a vector in R, and collect these into a single data frame.**

```{r}
P=c(0.27,0.21,0.34,0.32,0.24,0.29,0.18,0.20,0.29)
GR=c(1.79,1.01,1.71,1.89,1.41,1.15,1.13,1.20,1.10)
ConcVRate=data.frame('Peroxide'=P,'Growth.rate'=GR)
```

**Question 6. Create a scatter plot of fungal growth rate versus peroxide concentration from the newly created data frame.**

```{r}
plot(ConcVRate$Peroxide,ConcVRate$Growth.rate,xlab="Hydrogen peroxide concentration (µmoles/g)",ylab="Fungal growth rate (cm/day)",main="Leaf [H2O2] vs fungal growth rate")
```

**Question 7. How does the growth rate of the fungus change with increasing levels of hydrogen peroxide, and does this agree or disagree with the hypothesis?**

The growth rate of fungus does not decrease with increased hydrogen peroxide concentration in the leaves, as per the hypothesis. The data does not support the hypothesis.

### Part II - Descriptive statistics and graphing

**Question 8. Recreate this table in Excel and save as a CSV file. Read the data into R using the read.csv function.**

```{r}
turtles=read.csv("turtles.csv")
```

Note: the *file.choose* function created an error when trying to knit.

**Question 9. Use the table function to create a contingency table showing the number of aborted and successful nesting attempts for each species.**

```{r}
table(turtles)
```

**Question 10. Make and label a grouped barplot that represents the nesting patterns for all three turtle species.**

```{r}
turtles.df=table(turtles)
barplot(turtles.df,names.arg=turtles$species,beside=TRUE,main='Nesting patterns of turtle species',xlab='Turtle species',ylab='Frequency',col=c('firebrick','darkolivegreen'))
```

**Question 11. Determine the mean and standard deviation of successful nesting events among all marine turtles.**

```{r}
success=turtles.df[2,]
mean.success=mean(success)
std.dev.success=sd(success)
```

The mean number of successful nesting events among all marine turtles is `r mean.success` (standard deviation is `r std.dev.success`).

**Question 12. Calculate the proportion of failed nesting attempts for each species. Based on this, which species would you choose as needing the greatest conservation effort?**

```{r}
aborted=turtles.df[1,]
failed.olive=aborted[3]
success.olive=success[3]
prop.failed.olive=failed.olive/success.olive

failed.lb=aborted[2]
success.lb=success[2]
prop.failed.lb=failed.lb/success.lb

failed.hb=aborted[1]
success.hb=success[1]
prop.failed.hb=failed.hb/success.hb
```

The proportion of failed nesting attempts for Olive Ridley, Leatherback and Hawksbill turtles are `r prop.failed.olive`, `r prop.failed.lb`, and `r prop.failed.hb` respectively. Based on this data, Hawksbill turtles have the highest proportion of failed nesting attempts and may require the greatest conservation effort. 

**Question 13. Enter turtle egg data into Excel and import the file into R.**

```{r}
turtles2=read.csv("turtles2.csv")
```

Note: the *file.choose* function created an error when trying to knit.

**Question 14. Create a boxplot comparing different turtle species and the number of eggs found in the nest.**

```{r}
boxplot(turtles2$Olive.Ridley,turtles2$Hawksbill,turtles2$Leatherback,names=c("Olive Ridley","Hawksbill","Leatherback"),main="Number of eggs per nesting site by turtle species",ylab="Number of eggs per nest",xlab="Species",col=c('aquamarine3','plum3','lightgoldenrod3'))
```

**Question 15. Create a scatter plot of the number of Leatherback turtle eggs (y-axis) against the number of Hawksbill eggs (x-axis) across all beach sections.**

```{r}
plot(turtles2$Hawksbill,turtles2$Leatherback,xlab='Number of Hawksbill eggs per nest',ylab='Number of Leatherback eggs per nest',main='Comparing the number of Leatherback and Hawksbill eggs per nest')
```

